---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "150%"
)
library(cowplot)
```
# betaC

The r package "betaC" provides code for the calculation of beta_C, a metric to quantify the non-random component in beta-diversity for a given coverage.

## Installation

You can install the development version of betaC from [GitHub](https://github.com/T-Engel/betaC) with:

``` r
devtools::install_github("T-Engel/betaC")
```
Please, also install the package "vegan" from CRAN. Furthermore, "tidyverse" is recommended but the main functions will work without it. 

## Quick guide
If you are familiar with the concept of beta_C and you just want to have the code, here is all you need to know. If you're unfamiliar with the method you might want to read the paper and go through the basic examples that follow below.

* The main function of this package is `beta_C(x, C)` It's first argument `x`takes a site-by-species abundance matrix as a matrix object or data frame (sites=rows, species= rows). The second argument `C` is target coverage used for standardization. The function returns beta_C for the target coverage as a numeric value. Type `?beta_C` to see the documentation.

* The other important function is `C_target(x)`. Its only argument `x` is a site-by-species abundance matrix. It returns the maximum possible coverage value that can be used to calculate beta_C for the community matrix `x`. When comparing beta_C across multiple communities it is recommended to choose the `C` argument  in `beta_C(x, C)`such that it corresponds to the smallest `C_target()` output of all the communities. Type `?C_target to see the documentation.

## Example: Beta-diversity of the BCI dataset from vegan

Let's look at the beta_diversity of the `BCI` dataset that comes with the package vegan. It's a site by species abundance matrix with 50 plots and 225 species from the Barro Colorado Island in Panama. If you want to know more about the dataset type `?BCI` after loading the the vegan package. We are intereted in the beta-diversity of the island because it tells us something about the spatial structure species diversity. First, let's calculate Whittaker's multiplicative beta-diverity as
$$\beta=\frac{\gamma}{\overline{\alpha}}$$,
where $\gamma$ is the gamma species richness (i.e. all plots combined) and $\overline{\alpha}$ is the alpha species richness (i.e. the average plot richness). 

```{r example}
library(betaC)
library(vegan)
data(BCI)

# Whittakers multiplicative beta-diversity 
gamma=specnumber(colSums(BCI))
alpha= mean(specnumber(BCI))
beta_BCI=gamma/alpha
beta_BCI

```

The samples have a beta-diversity of `r round(beta_BCI,2)` The unit of this value is "effective number of distinct sampling units". This means that it takes `r round(beta_BCI,2)` hypothetical plots with complete species turnover to produce the same beta-diversity that we observe in this data-set. Note that the maximum possible value of beta is 50 here. In that case all 50 plots would be completely unique in their species identities. The minimum value of beta is 1 in which case all the plots share the same species.  However, this assumes that the sample size is big enough to randomly sample all the species in the area - in other words that the samples are complete. While this may be true on the gamma scale, where we have a really large sample size, the alpha scale has a way smaller number of individuals. Remember that the gamma scale has all the individuals of the 50 subplots combined, while the average plot has only a 50th of this sample size. Let's have a look at this difference:

```{r}
N_alpha= mean(rowSums(BCI))
N_gamma= sum(BCI)
barplot(c("gamma"=N_gamma,"alpha"= N_alpha), ylab = "Number of individuals", xlab = "Scale")
```

To controll for this vast difference in sample size between the scales we use individual based rarefaction (IBR). IBR lets you ask the question "How many species can I expect to find if I sampled *n* Individuals rather than the actual sample size of *N*?" This enables us to rescale our gamma diversity estimate to the sample size observed at the alpha scale. The relationship between the rarefied richness and the sample size *n* used for rarefaction is called a individual based rarefaction curve.


To illustrate this let's have a look at the two-scale rarefaction curve.

```{r  message=FALSE, warning=FALSE}
library(tidyverse)

# alpha 
alphas_curves<-rarecurve(BCI)
N_min= min(rowSums(BCI))
mean_alpha_curve= sapply(alphas_curves,function(x) return(as.numeric(x[1:N_min])) ) %>% rowMeans()

#gamma
gamma_curve= as.numeric(rarefy(colSums(BCI),sample = 1:N_gamma))

# plot them
plot(gamma_curve, xlab = "Number of individuals", ylab= "Rarefied richness", type = "l", lwd=2)
lines(mean_alpha_curve, col =3, lwd=2)
```




## Functions
The main function `beta_C` uses the following  helper functions to jump between sample size and coverage. Both have been adapted from iNEXT.

- `Chat`
- `invChat`

Furthermore, there is a function that helps to determine the target coverage value:

- `C_target`


<!-- badges: start -->
  [![Travis build status](https://travis-ci.org/T-Engel/betaC.svg?branch=master)](https://travis-ci.org/T-Engel/betaC)
  <!-- badges: end -->
