---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "150%"
)
library(cowplot)
```
# betaC

The r package "betaC" provides code for the calculation of beta_C, a metric to quantify the non-random component in beta-diversity for a given coverage. 

## Installation

You can install the development version of betaC from [GitHub](https://github.com/T-Engel/betaC) with:

``` r
devtools::install_github("T-Engel/betaC")
```

Please, also install the package "vegan" from CRAN. Furthermore, "tidyverse" is recommended but the main functions will work without it. 

## Example: Beta-diversity of the BCI dataset from vegan

Let's look at the beta_diversity of the BCI dataset that comes with the package vegan. First we calculate Whittaker's multiplicative beta-diverity and than we standardize it to a sample coverage of 50% using `beta_C`.

```{r example}
library(betaC)
library(vegan)
data(BCI)

# Multiplicative beta-diversity 
gamma=specnumber(colSums(BCI))
alpha= mean(specnumber(BCI))
beta_BCI=gamma/alpha
beta_BCI

# beta_C for 50% coverage
beta_C_BCI<-beta_C(BCI,0.5)
beta_C_BCI
```

The samples have a beta-diversity of `r round(beta_BCI,2)` but most of this sample differention is due to a sampling effect. If we standardize to the number of individuals that corresponds to 50 % coverage at the gamma scale we find that the partition of the non-random component in beta-diversity is only ca. `r round(beta_C_BCI,2)`

To illustrate this let's have a look at the two-scale rarefaction curve.

```{r  message=FALSE, warning=FALSE}
library(tidyverse)

# calculate rarefaction curves
dat<-betaC:::rarefy_long(as.matrix(BCI))

# plot them
dat %>% filter(type== "major") %>% ggplot(aes(N,S_n, col= Curve))+ geom_line(size=1)+ geom_hline(yintercept =  alpha)+
geom_hline(yintercept =  gamma)+ geom_vline(xintercept =  invChat(colSums(BCI), 0.5), linetype= "dashed")
```

The horizontal black lines indicate the observed species richness at alpha and gamma scales. The rarefaction curves fall very closely on top of each other indicating that there is not a strong signal of spatial structure. The observed beta-diversity is mostly caused by difference in sample size. Beta_C of 50% is calculated at a sample size of `r round(invChat(colSums(BCI), 0.5))` individuals (dashed vertical line).

For this example we chose the 50% arbitrarily. We can also examine the entire scaling relationship of the nonrandom component in beta-diversity. To do that we calculate beta_Sn for every N and the corresponding coverages at the gamma scale. This can be done using `beta_C_curve`. 

```{r}
BCI_curve<-beta_C_curve(BCI)

head(BCI_curve)
```

And here are the plots:

```{r}

N_plot <-
    BCI_curve %>% ggplot(aes(N, beta_Sn)) + geom_line(size = 1) + geom_rug() +
    geom_hline(yintercept = beta_C_BCI,
               col = "red",
               linetype = "dotted") + geom_vline(
                   xintercept = invChat(colSums(BCI), 0.5),
                   col = "red",
                   linetype = "dotted"
               )
C_plot <-
    BCI_curve %>% ggplot(aes(C, beta_Sn)) + geom_line(size = 1) + geom_rug() +
    geom_hline(yintercept = beta_C_BCI,
               col = "red",
               linetype = "dotted") + geom_vline(
                   xintercept =  0.5,
                   col = "red",
                   linetype = "dotted"
               )
plot_grid(N_plot, C_plot, align = "v")
```

Note that both graphs show the same data on the y axis. The value corresponding to 50% that we calculated previously is indicated by the dotted red lines. The coverage/ sample size value used for the standarisation stronly influences the value of beta_C. Therefore it should always be reported and compared with respect to the underlying coverage.

## Functions
The main function `beta_C` uses the following  helper functions to jump between sample size and coverage. Both have been adapted from iNEXT.

- `Chat`
- `invChat`

Furthermore, there is a function that helps to determine the target coverage value:

- `C_target`


